Efficiently Handling Guards in the Low Level Design of RPython's Tracing JIT

### 2012-08-07 talk/vmil2012/paper.tex
(cfbolz, bivab) write section about RPython's tracing JIT

\subsection{RPython's Meta-Tracing JIT Compilers}
\label{sub:tracing}

 * Tracing JITs
 * Mention SSA
 * JIT Compiler
   * describe the tracing jit stuff in pypy
   * reference tracing the meta level paper for a high level description of what the JIT does
   * JIT Architecture
   * Explain the aspects of tracing and optimization

>>>>

\subsection{RPython's Tracing JIT Compilers}
\label{sub:tracing}
Tracing JITs are a technique of just-in-time compilers that generate code by
observing the execution of a program. VMs using tracing JITs are typically
mixed mode execution environments containing also an interpreter. The
interpreter profiles the executed program and selects frequently executed code
paths to be compiled to machine code. After profiling identified an interesting
path, tracing is started, recording all operations that are executed on this
path. Like in most compilers tracing JITs use an intermediate representation
to store the recorded operations, which is typically in SSA form\todo{some ssa
reference}. Since tracing follows actual execution the code that is recorded
represents only one possible path through the control flow graph. Points of
divergence from the recorded path are marked with special operations called
\emph{guards}, these operations ensure that assumptions valid during the
tracing phase are still valid when the code has been compiled and is executed.
After a trace has been recorded it is optimized and then compiled to platform
specific machine code.

When the check of a guard fails, the execution of the machine code must be
stopped and the control is returned to the interpreter, after the interpreter's
state has been restored. If a particular guard fails often a new trace is
recorded starting from the guard. We will refer to this kind of trace as a
\emph{bridge}. Once a bridge has been traced it is attached to the
corresponding guard by patching the machine code. The next time the guard fails
the bridge will be executed instead of leaving the machine code.

RPython provides a tracing JIT that can be reused for a number of language
implementations. This is possible, because it traces the execution of the
language interpreter instead of tracing the user program directly. This
approach is called \emph{meta-tracing}. For the purpose of this paper the fact
that RPython's tracing JIT is a meta-tracing JIT can be ignored.
